---
title: "Complex Production Function"
author: "EB"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE,
  warning = FALSE
  )
```

## Reproducing CD production function from the slides

Estimate a complex Cobb-Douglas production with all relevant inputs that are present in the data.

$$
\begin{aligned}
\ln (\text{output}) & = \alpha + \beta_1 \text{status : share} + \beta_2 \text{status : mixed} \\
& + \beta_3 \ln (\text{land}) + \beta_4 \ln (\text{labor_family}) + \beta_5 \ln (\text{labor_hired}) \\
& + \beta_6 \ln (\text{seed}) + \beta_7 \ln (\text{urea}) + \beta_8 \ln (\text{phosphate}) + \beta_9 \ln (\text{pesticide})  + \\ 
& + \beta_{10} \text{region} + e
\end{aligned}
$$

As data contains observations with values of zero. Replace all zeros with a small number (0.00001, for example) before performing log transformation.

### Setup and data loading

```{r}
library(tidyverse)
library(GGally)
library(modelsummary)
library(performance)
library(lmtest)

farms <- read_csv("rice.csv")
```

### Exploratory data analysis

```{r}
farms %>% datasummary_skim()
```

### Estimating regressions

#### Bonus reverse dummy variables

To resolve log of zero problem, you may introduce reverse dummy variables. These are additional regressors, one per each independent variable transformed with a log. Here is the algorithm of reverse dummy variables introduction:

-   Reverse dummy variables are introduces one by one. One reverse dummy variable per an independent variable transformed with log.
-   Reverse dummy variables contains zero values.
-   When corresponding regressor is zero ($log(0) = -\infty$), reverse dummy variable is one.
-   Then, take log of the main regressors.
-   Replace any negative infinite values of the logged variables with zero.
-   Run regression with all regressors and reverse dummy variables.

#### Dealing with log(0)

```{r}
fit_cd_1 <- lm(output ~ 1, data = farm_dta)
```

### Validating assumptions

#### Linearity

#### Multicollinearity

#### Homoscedasticity

### Reporting regression results

-   Produce robust standard errors

```{r}
summary(fit2_cd)
```
