---
title: "Panel regression production function"
author: "EB"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE,
  warning = FALSE
  )
```

# Introduction and the research question

Are larger farms more productive if compared to the smaller once?

Cobb-Douglas production functional is suggested by the theory to explore the size-productivity relationship.

$$
\ln (\text{output}_{it}) = A_0 
+ \beta_1 \cdot \ln (\text{land}_{it}) 
+ \beta_2 \cdot \ln (\text{labor hired}_{it}) \\
+ \beta_3 \cdot \ln (\text{labor family}_{it}) \\
+ \beta_4 \cdot \ln (\text{seed}_{it}) 
+ \beta_5 \cdot \ln (\text{urea}_{it}) \\
+ \beta_6 \cdot \ln (\text{pesticide}_{it})
+ \beta_7 \cdot \text{variety(high)}_{it} \\
+ \beta_8 \cdot \text{variety(trad)}_{it}
+ \beta_{9-12} \cdot \text{Region}_{i}
+ e_{it}
$$

Variables used are:

-   `id` the farm identifier
-   `time` the growing season
-   `size` the total area cultivated with rice, measured in hectares
-   `status` land status, on of 'owner' (non sharecroppers, owner operators or leaseholders or both), 'share' (sharecroppers), 'mixed' (mixed of the two previous status)
-   `varieties` one of 'trad' (traditional varieties), 'high' (high yielding varieties) and 'mixed' (mixed varieties)
-   `bimas` is an intensification program ; one of 'no' (non-bimas farmer), 'yes' (bimas farmer) or 'mixed' (part but not all of farmer's land was registered to be in the bimas program)
-   `seed` seed in kilogram
-   `urea` urea in kilogram
-   `phosphate` phosphate in kilogram
-   `pesticide` pesticide cost in Rupiah
-   `hiredlabor` hired labor in hours
-   `famlabor` family labor in hours
-   `totlabor` total labor (excluding harvest labor)
-   `region` one of wargabinangun, langan, gunungwangi, malausma, sukaambit, ciwangi
-   `price` price of rough rice in Rupiah per kg

Write down the regression equation that could be estimated using the fixed effect panel regression. Make sure that time invariant variables are not appearing in the equation.

## Write down the regression equation

$$
\ln (\text{output}_{it}) = 
$$

# Data exploration

To explore the data. Use `farm_panel.csv`.

## Load the data

-   Load the data.
-   Compute rice yields.
-   Compute the value of output (output times price).

```{r}

```

## Summary statistics for key variables

Report summary statistics for non-logged variables.

```{r}

```

Report mean, SD, median, min, max statistics for farm size, hired and family labor, output and yields by region.

```{r}

```

## Explore relationship farm size and rice yields

1.  Build a scatter plot;
2.  Add colors for `varieties`;
3.  Make different plots by region;
4.  Ensure that all labels are specified properly.

```{r}

```

# Estimate regressions models

## Pooled

```{r}

```

### Test key assumptions

```{r}

```

## FE (individual and two-ways)

```{r}

```

### Perform model selection tests

```{r}

```

## RE

```{r}

```

### Perform model selection tests

```{r}

```

## Report results for the pooled and selected model

-   Briefly explain why you have selected this model.
-   Make sure to perform the robust inference.
-   Describe the results

```{r}

```

## Explore relationship between farm size and fixed effects

Extract the fixed effects from the selected model

```{r}

```

Compute average farm size

```{r}

```

Plot the relationship

```{r}

```

Rerun fixed effect regression using the value of output as a dependent variable. Extract and plot the fixed effects vs farm size. Are the relationship between efficiency and size different between two plots?

```{r}

```
