---
title: "Union membership and wage relationship"
author: "EB"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE,
  warning = FALSE
  )
```

# Introduction and the research question

Does the collective bargaining (union membership) has any effect on wages?

Following functional form is suggested by the theory and will be estimated:

$$
\log \text{Wage}_{it} = \beta_0 + \beta_1 \cdot \text{Union}_{it} + \beta_2 \cdot {X_{it}} + \epsilon_{it},
$$

where ${X_{it}}$ is the vector of control variables such as: age, age squared, experience, tenure, education and gender.

As data that we are using contains some panel structure, write down the regression equation that could be estimated using the fixed effect panel regression:

## Write down the regression equation

# Data exploration

To explore this relationship, we rely on the data `wage_unon_panel_v2.csv` (different from the one presented on the slides).

## Load the data

```{r}

```

## Summary statistics for key variables

```{r}

```

## Explore relationship between wage and experience

1.  Build a scatter plot;
2.  Add colors for union membership;
3.  Make different plots year (at least first and last year);
4.  Ensure that all titles are specified properly.

```{r}

```

# Estimate regressions models

## Pooled

```{r}

```

### Test key assumptions

```{r}

```

## FE (individual and two-ways)

```{r}

```

### Perform model selection tests

```{r}

```

## RE

```{r}

```

### Perform model selection tests

```{r}

```

## Report results for the pooled and selected model

-   Briefly explain why you have selected this model.
-   Make sure to perform the robust inference.
-   Describe the results

```{r}

```
