---
title: "Wine regression"
author: "EB"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Using data published in 

-    Orley Ashenfelter , David Ashmore & Robert Lalonde (1995) Bordeaux Wine Vintage Quality and the Weather, CHANCE, 8:4, 7-14, DOI: 10.1080/09332480.1995.10542468

we would like to estimate what are the determinants of wine prices.

Following variables are present in the data: 

-   `Year`: year in which grapes were harvested to make wine.
-   `Price`: average market price for Bordeaux vintages according to a series of auctions. The price is relative to the price of the 1961 vintage, regarded as the best one ever recorded.
-   `WinterRain`: winter rainfall (in mm).
-   `AGST`: Average Growing Season Temperature (in Celsius degrees).
-   `HarvestRain`: harvest rainfall (in mm).
-   `Age`: age of the wine, measured in 1983 as the number of years stored in a cask.
-   `FrancePop`: population of France at Year (in thousands).

## Setup

Load (install) packages below and load the data. Packages are: 

-   `tidyverse`
-   `GGally`
-   `modelsummary`
-   `correlation`
-   `performance`

```{r}
# library()
# library()
# library()
# library()
# library()
# wine_dta <- 
```

## The data

```{r}
glimpse(wine_dta)
```

### Descreptive statistics

```{r}
```


### Correlation matrix

```{r}
```


### Visual inspection

- Remove columns year and population and plot data pairs

```{r message=FALSE}
```

## Regression analysis

### Molde 1: level-level

$\text{Price} = \beta_0 + \beta_1 \text{Age} + \beta_2 \text{AGST} + \beta_3 \text{WinterRain} + \beta_4 \text{HarvestRain} +e$

```{r}
```


Fancy layout

```{r}
```


#### Checking the linearity assumption

```{r}
```
### Model 2 Log-level

$\log (\text{Price}) = \beta_0 + \beta_1 \text{Age} + \beta_2 \text{AGST} + \beta_3 \text{WinterRain} + \beta_4 \text{HarvestRain} +e$

```{r}
```

#### Checking linearity again

```{r}
```

### Model 3 Log-level with age only

Estimate model: $\log (\text{Price}) = \beta_0 + \beta_1 \text{Age} +e$

```{r}

```


## Reporting regressions results sude-by-side

```{r}
```


## Solutions

```{r eval=FALSE}
```

