---
title: "Wage - education relationship"
author: "EB"
date: "`r Sys.Date()`"
output: html_document
---

## Goals

-   Fit multiple linear regression;
-   Display and interpret regression summary;
-   Compute fitted values and residuals;

## Data

We use data from Wooldridge 2015 on wage determinants. Variables present are:

-   $wage$ - hourly earnings in USD;
-   $educ$ - years of education;
-   $exper$ - years of experiences;
-   $black$ - dummy variable representing individuals which are not Caucasian;
-   $female$ - dummy variable representing female;

See also for more information: Blackburn, M. and Neumark, D. (1992) Unobserved Ability, Efficiency Wages, and Interindustry Wage Differentials. The Quarterly Journal of Economics , Vol. 107, No. 4. Oxford University Press (OUP). p. 1421-1436

Loading (installing) libraries:

```{r}

```


```{r}
# wage_dta <- read_csv(______________)
```

## Ex. 1. Glimpse at the data

```{r}
# __________(_____)
```

Compute a new variables `caucasian` that takes values "yes" and "no" for each person depending on the values of the variable `black`.

```{r}
# wage_dta <-
#   _____ %>%
#   mutate(_____ = ifelse(black, "no", "yes"),
#          caucasian = as.factor(caucasian))
# glimpse(_____)
```

### Ex. 1.2. Descriptive statistics

Use function `datasummary_skim()` from the package `modelsummary`.

```{r}
# library(_________)
# _________(wage_dta)
```

### Ex. 2.3. Visual data inspection

We inspect data visually by building a grid of scatter plots using `GGally::ggpairs()`

```{r}
# library(_________)
# _________(wage_dta)
```

Add parameter `aes(colour = caucasian)` to `ggpairs()`.

What is the difference compare to the previous plot?

```{r}
# ggpairs(_________, _________)
```

## Ex. 2. Fit regressions

Dependent variable is `wage`, independent are: `educ`, `exper` and `black`

```{r}
# fit1 <- lm(____ ~ ____ + ____ + ____, data = ____)
# fit1
```

### Ex. 2.1. Print regression results

```{r}
# fit1
```

## Ex. 3 Regression summary

```{r}
# summary(______)
```

Interpret the coefficients:

### Ex. 2.1 Regression summary using `parameters::parameters()`

Check help on `?parameters::parameters`

```{r}
# library(________)
# ________(________)
```

### Ex. 2.2. Goodness of fit using `performance::performance()`

Check help on `?performance::performance`

```{r}
# library(________)
# performance(________)
```

## Ex. 4. Fitted

```{r}
# fitted_vector <- fitted(________)
# fitted_vector[1:20]
```

## Ex. 5. Residuals

```{r}
# resid_vector <- resid(________)
# resid_vector[1:20]
```

## Ex. 6. Residuals vs fitted plot

Use function `plot()`. Put fitted as an `x` argument and residuals as `y`.

```{r}
# plot(x = _____, ___ = ________)
```

### Ex. 6.1 Residuals vs fitted using `check_model` from `performance` package

See: `?performance::check_model`

```{r}
# check_model(fit1, check = "linearity")
```

## Ex. 7 Predicting values using regression

We can compute predicted values based on the regression results for arbitrary values of $X$. See: `?predict`

First, we need to make a table the $X$ variables for which we want compute the predicted value.

-   Variables names in table should match the variables in the regression.

Let us predict wage for a person with 0 years of education, 0 experience, when he/she is white.

```{r}
# pred1 <-
#   tibble(educ = 0,
#          exper = 0,
#          black = 0)
# predict(fit1, pred1)
```

The same when the person is black.

```{r}
# pred2 <-
#   tibble(educ = c(0, 0),
#          exper = c(0, 0),
#          black = c(0, 1)
#   )
# predict(fit1, pred2)
```

### Ex 7.1 Writing predicted values to a data frame

Use mutate to mutate new variable in a data frame:

```{r}
# pred3 <-
#   tibble(educ = c(0, 0, 10),
#          exper = c(0, 0, 25),
#          black = c(0, 1, 1)
#   )
# 
# pred3 %>% 
#   mutate(predicted = predict(fit1, pred3))
```

### Ex 7.2 Visual inspection of predicted values

We can use function `ggeffects::ggpredict()`, check help.

```{r}
# library(ggeffects)
# ggpredict(fit1, term = "educ")
```

We can also plot this effects using `plot` after `ggpredict`


```{r}
# ggpredict(fit1, term = "educ") %>% ________()
```

Compute predicted values for a different independent variable and a dummy variable

```{r}
# ggpredict(fit1, term = c("________", "________")) 
# ________(________, term = c("________", "________")) %>% plot()
```

