---
title: "Hedonic regression"
author: "EB"
date: "`r Sys.Date()`"
output: html_document
---

# Goals:

-   Get a brief introduction to the Hedonic Prices model;
-   Practice fitting linear regression in R;
-   Test linearity assumption: visually and using a statistical tests;
-   Practice linear transformation;
-   Exerciser coefficients interpretation;

# Problem

We would like to assess the effect of the "Conservation Reserve Program" (CPR) on the agricultural land prices in Minnesota in 2002-2011.

Conservation Reserve Program

-   is a subsidy
-   obligates farms NOT TO GROW ANY CROPS on the enrolled land
-   pays monetary compensation in exchange;

As this is a risk less income, it may increase the land price. However, if the land is productive, such set-aside measure may reduce farmland price.

We would like to estimate a following equation:

$$
\text{acrePrice} = \beta_0 +  \beta_1 \text{crpPct} + \beta_2 \text{acres} + \beta_3 \text{region} + \beta_4 \text{year} + \beta_5 \text{tillable} + \beta_6  \text{productivity} + e
$$

-   `acrePrice` - sale price in dollars per acre;
-   `acres` - size of the farm in acres;
-   `region` - region in the state Minnesota;
-   `year` - year of the land sales translation;
-   `crpPct` - the percentage of all farm acres enrolled in CRP;
-   `tillable` - percentage of farm acreage that is rated arable by the assessor;
-   `productivity` - average agronomic productivity scaled 1 to 100, with larger numbers for more productive land;

# Exercise

## 0. Setup

We will need following packages:

-   `tidyverse`
-   `readxl`
-   `GGally`
-   `modelsummary`
-   `correlation`
-   `performance
-   `car`
-   `lmtest`
-   `sandwich`

```{r}

```


## 1. Load the data

```{r}
# dta <- 
# glimpse(dta)
```

## 2. Draw box-plots of `acrePrice` versus `year`

Modify variable `year` creating a factor based on `year`: use function `as.factor()`;

Play with the scale transformation of `acrePrice` to visualize data more meaningful:

-   check help for `ggplot2::scale_*_continuous(trans = "_______")`
-   check help for `ggplot2::scale_*_log10()`

```{r}
# dta %>% 
#   _________(________ = __________(________)) %>% 
#   ggplot() + 
#   aes(___ = ____, ___ = ____) + 
#   geom_________()
```

## 3. Convert monetary values from current to constant prices

Create a table with GDP deflator rates

```{r}
defl_dta <-
  tibble(
    year = 2002:2011,
    defl = c(77.47, 78.91, 81.03, 83.56, 86.09, 
             88.4, 90.12, 90.8, 91.86, 93.78)
  )
glimpse(defl_dta)
```

Join `defl_dta` to the `dta` data and create new object `dta_const`:

-   see help on joining data [here](https://rstudio-education.github.io/tidyverse-cookbook/transform-tables.html#joins);
-   convert `acrePrice` to the constant prices.
-   remove deflator variable;
-   convert variable year to a factor;

```{r}
# dta_const <- 
#   _________ %>% 
#   left_join(_________, by = "_________") %>% 
#   _________(_________ = _________ * (_________),
#             year = _________(_________)) %>% 
#   _________(- _________)
# glimpse(dta_const)
```

HW. Built a box plot out of it with log transformation.

```{r}
#
```

## 4. Produce summary statistics

Use any function of your liking. Maybe data summary skim from `modelsummary` package.

```{r}
#
```

## 5. Produce a correlation matrix

Use: `correlation::correlation()`

```{r}
# ___________ %>% 
#   ___________() %>% 
#   summary()
```

## 6. Fit regression and summarize the results

use `log(acrePrice)` as a dependent variable

```{r}
# fit1 <- lm(
#   _____ ~ _____,
#   data = _____
# )
# summary(fit1)
```

Use `performance` package to summarize the regression GOF and `modelsummary` to print the results.

```{r}
# performance(_____)
```

```{r}

```


### Interpret the regression results

```{r}
# modelsummary(fit1)
```

## 7. Check linearity visually

```{r}
# check_model(______, check = ______)
```

## 11. Check multicollinearity

use `vif()` from `car`.

```{r}
# ______(______)
```

## 12. Check homoscedasticity visually

```{r}
# check_model(______, check = c("linearity", "homogeneity"))
```

## 13. Check homoscedasticity using statistical tests

```{r}
library(lmtest)
# ______(______)
```

## 14. Correct standard errors and interpret the results

```{r}
library(sandwich)
# parameters(_____, vcov = "HC3")
```
