---
title: "Interaction terms"
author: "EB"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE,
  warning = FALSE
  )
```

Our goal is to improve on the regression covered in the class and resolve the non-linearity problem that is present there. Then we will proceed to interpreting the interaction terms.

# Loading and exploring data

```{r}
library(tidyverse)
library(readxl)
library(modelsummary)
library(ggeffects)
library(marginaleffects)
library(performance)
library(GGally)
library(car)

## 1. Load the data
dta <-
  read_excel("land-prices.xlsx") %>%
  mutate(improvements  = as.numeric(improvements),
         development = as.factor(as.integer(improvements > 25)),
         productivity  = as.numeric(productivity),
         tillable  = as.numeric(tillable),
         year = as.factor(year)) %>% 
  select(acrePrice, crpPct, acres, improvements, development, region, year)
```

## Summary statistics

```{r}
```

## Base linear regression

```{r}
```

Calculating marginal effects manually

```{r}
```

Calculating standard errors using delta method

```{r}
```

Same using `slopes` function:

```{r}
```

Displaying model results and marginal effects.

```{r}
```

Marginal effects at an arbitrary point of the interaction term.

```{r}
```

## Log-log regression

```{r}
```

Marginal effect of the $\log x$ on the $\log y$ (elasticity): $\frac{\partial \log y}{\partial \log x}$.

Using `margins` package:

```{r}
library(margins)
```

ME with slopes

```{r}
```

ME manually

```{r}
```

Me at specific values of interaction terms

```{r}
```

Marginal effect of $x$ on $y$ level-level equivalent: $\frac{\partial y}{\partial x}$.

```{r}
```

Me at specific values of interaction terms.

```{r}
```

## Reporting and interpreting the results

```{r}
```
