---
title: "Interaction terms"
author: "EB"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE,
  warning = FALSE
  )
```

Our goal is to improve on the regression covered in the class and resolve the non-linearity problem that is present there. Then we will proceed to interpreting the interaction terms.

# Loading and exploring data

```{r}
library(tidyverse)
library(readxl)
library(modelsummary)
library(ggeffects)
library(marginaleffects)
library(performance)
library(GGally)
library(car)

## 1. Load the data
dta <-
  read_excel("land-prices.xlsx") %>%
  mutate(improvements  = as.numeric(improvements),
         development = as.factor(as.integer(improvements > 25)),
         productivity  = as.numeric(productivity),
         tillable  = as.numeric(tillable),
         year = as.factor(year)) %>% 
  select(acrePrice, crpPct, acres, improvements, development, region, year)

```

## Summary statistics

```{r}
n_nonmis <- function(x, ...) as.character(round(sum(!is.na(x)), 0))
datasummary(
    acrePrice + log(acrePrice) + 
      crpPct + 
      acres + log(acres) +
      improvements ~ 
      ((`N non-missing` = n_nonmis) + mean + sd + min + median + max) * 
      Arguments(na.rm = TRUE),
    data = dta
  )
```

## Base linear regression

```{r}
fit3 <- lm(acrePrice  ~ acres * improvements + crpPct + region + factor(year),
           data = dta)
check_model(fit3, check = c("linearity", "homogeneity"))
```

Calculating marginal effects manually

```{r}
coef(fit3)[["acres"]] + coef(fit3)[["acres:improvements"]] * mean(dta$improvements, na.rm = TRUE) 
coef(fit3)[["improvements"]] + coef(fit3)[["acres:improvements"]] * mean(dta$acres, na.rm = TRUE) 
```

Calculating standard errors using delta method

```{r}
library(car)
library(sandwich)
deltaMethod(fit3, g. = "acres + `acres:improvements` * 4.492922")

# With clustered SE
deltaMethod(fit3, g. = "acres + `acres:improvements` * 4.492922",
            vcov. = sandwich::vcovCL(fit3, cluster = ~region))

deltaMethod(fit3, g. = "improvements + `acres:improvements` * 112.6855")

# With clustered SE
deltaMethod(fit3, g. = "improvements + `acres:improvements` * 112.6855",
            vcov. = sandwich::vcovCL(fit3, cluster = ~region))
```

Same using `slopes` function:

```{r}
slopes(fit3, variables = c("acres", "improvements"), 
       slope = "dydx", 
       newdata = "mean")
slopes(fit3, variables = c("acres", "improvements"), 
       slope = "dydx", newdata = "mean", 
       vcov = ~region)
```

Displaying model results and marginal effects.

```{r}
modelsummary(
  list(
    `Regression as is` = fit3,
    `Marginal Effects` =
      slopes(fit3, variables = c("acres", "improvements", "crpPct"), 
             slope = "dydx", newdata = "mean", 
             vcov = ~region)
  ), 
  estimate = "{estimate}{stars} ({std.error})",
  fmt = 1,
  statistic = NULL,
  vcov = list(~region, NULL), 
  coef_omit = "reg|year",
  coef_map = c("acres" = "Area, ha",
               "improvements" = "Improvements, %",
               "crpPct" = "Subsidy, ha",
               "acres:improvements" = "Area * Improvements"
               ),
  gof_omit = "AIC|BIC|F|Log|RMS"
  )
```

Marginal effects at an arbitrary point of the interaction term.

```{r}
slopes(fit3, 
       variables = c("improvements"), 
       slope = "dydx", 
       vcov = ~ region, 
       newdata = datagrid(acres = fivenum(dta$acres)))
```

## Log-log regression

```{r}
fit4a <- 
  lm(lprice  ~ larea * improvements + crpPct + region + factor(year),
     data = dta %>% mutate(larea = log(acres), lprice = log(acrePrice)))
```

Marginal effect of the $\log x$ on the $\log y$ (elasticity): $\frac{\partial \log y}{\partial \log x}$.

Using `margins` package:

```{r}
library(margins)
summary(margins(fit4a,
                variables  = c("larea", "improvements"),
                vcov = sandwich::vcovCL(fit4a, cluster = ~region)))
```

ME with slopes

```{r}
slopes(fit4a,
       variables = c("larea", "improvements"), 
       slope = "dydx", 
       vcov = ~ region, 
       newdata = "mean")
```

ME manually

```{r}
mean(model.frame(fit4a)$improvements)
deltaMethod(fit4a, g. = "larea + `larea:improvements` * 4.492922",
            vcov. = sandwich::vcovCL(fit4a, cluster = ~region))

mean(model.frame(fit4a)$larea)
deltaMethod(
  fit4a, g. = "improvements + `larea:improvements` * 4.419874",
  vcov. = sandwich::vcovCL(fit4a, cluster = ~region))
```

Me at specific values of interaction terms

```{r}
# ME with slopes
slopes(
  fit4a,
  variables = c("improvements"),
  slope = "dydx",
  vcov = ~ region,
  newdata = datagrid(larea = log(fivenum(dta$acres)))
)

slopes(
  fit4a,
  variables = c("larea"),
  slope = "dydx",
  vcov = ~ region,
  newdata = datagrid(improvements = c(1, 25, 50, 75, 100))
)
```


Marginal effect of $x$ on $y$ level-level equivalent: $\frac{\partial y}{\partial x}$.

```{r}
fit4 <- 
  lm(log(acrePrice) ~ log(acres) * improvements + crpPct + region + factor(year),
     data = dta)

# using margins package
summary(margins(fit4, variables  = c("acres", "improvements")))

# using Margins package
avg_slopes(
  fit4,
  variables = c("acres", "improvements"),
  slope = "dydx"
)
```


Me at specific values of interaction terms.

```{r}
# ME with slopes
slopes(
  fit4,
  variables = c("improvements"),
  slope = "dydx",
  vcov = ~ region,
  newdata = datagrid(acres = fivenum(dta$acres))
)
```

## Reporting and interpreting the results

```{r}
modelsummary(
  list(
    `Level-level` = fit3,
    `Log-log` = fit4a,
    `Level-level (ME)` = 
      slopes(fit3,
             variables = c("acres", "crpPct", "improvements"), 
             slope = "dydx", vcov = ~ region, newdata = "mean"),
    `Log-log (ME)` = 
      slopes(fit4a,
             variables = c("larea", "crpPct", "improvements"), 
             slope = "dydx", vcov = ~ region, newdata = "mean")
      
  ), 
  estimate = "{estimate}{stars} ({std.error})",
  # fmt = fmt_significant(2),
  statistic = NULL,
  vcov = list(~region, ~region, NULL, NULL), 
  coef_omit = "reg|year",
  coef_map = c("acres" = "Area, ha",
               "larea" = "Area, ha",
               "improvements" = "Improvements, %",
               "crpPct" = "Subsidy, ha",
               "acres:improvements" = "Area * Improvements"
               ),
  gof_omit = "AIC|BIC|F|Log|RMS"
  )
```

