---
title: "Reproducing the psychological costs of currency transition: evidence from the euro adoption papre."
author: "EB"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE,
  warning = FALSE
  )
```

## Introduction

The goal of this exercise is to reproduce results of the paper "Psychological costs of currency transition: evidence from the euro adoption" published by Otrachshenko et al in 2016 in the European Journal of Political Economy. This paper employed a difference in difference (DID/DD) approach for estimating a causal effect of transition to Euro in Slovakia on the individuals' satisfaction with their economy.

Authors estimate a simple model of individuals $i$ satisfaction with the economic situation in the sub-region $j$ and time period $t$ as a function of control variables $x^{'}_{ijt}$, regional fixed effect $\text{Region}^{'}_{j}$, indicator variable for Slovakia $\text{Slovak}_i$ and dummy variable that indicates that time period 2010 ($\text{y2010}_t$), after the transition to euro currency. 

$$
\text{Satisfaction}_{ijt} = \beta_0 + \beta_1 \text{y2010}_t + \beta_2  \text{Slovak}_i +  \alpha_{DD} (\text{Slovak}_i \cdot \text{y2010}_t) + \gamma x^{'}_{ijt}+ \eta \text{Region}^{'}_{j} + \epsilon_{ijt}
$$



-   What is our variable of interest?

-   What effect do we expect to find?

## Load and explore the data

```{r}
library(haven)
library(labelled)
library(tidyverse)
library(modelsummary)
library(car)
library(performance)
dta <- read_dta("otrachshenko2016.dta")
```

Note two new packages:

-   `labelled` Help here: [https://larmarange.github.io/labelled/](https://larmarange.github.io/labelled/)
-   `haven` Help here: [https://haven.tidyverse.org/index.html](https://haven.tidyverse.org/index.html)

Glimpse at the data and identify, what does each observations stands for?

```{r}
glimpse(dta)
```


## Explore the data

Using `count()` calculate how many observations are in each country/region/year.


```{r}
```

Plot a distribution of the satisfaction variable (add a country and year dimension).

```{r}
dta %>%
  ggplot() + 
  aes(x = stfeco, colour = as_factor(idcountry)) + 
  geom_density() + 
  scale_color_hue(NULL) +
  facet_wrap(. ~ year) +
  xlab("Satisfaction index")
```


Plot a box plot of the satisfaction by region and year.

```{r}
dta %>% 
  mutate(region = ifelse(idcountry == 1, str_c("SK ", idregion), str_c("CZ ", idregion)) %>% 
           factor()) %>% 
  ggplot() + 
  aes(x = as_factor(region), y = stfeco, 
      colour = as_factor(year)) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  scale_color_hue(NULL) +
  xlab("Region") + 
  ylab("Satisfaction index")
```


### Symmary statistics

Descriptive statistics by year of some key variables:

```{r}
nobs <- function(x, ...) as.character(round(sum(!is.na(x)), 0))

datasummary(
  stfeco + hhmmb + gndr + agea ~
    as_factor(year) * 
    (((Mean = mean) + (`Std. Dev.` = sd) + (`N obs.` = nobs)) * Arguments(na.rm = T)),
  data = dta %>%
    set_variable_labels(
      stfeco = "Satisfaction index (0-10)",
      hhmmb  = "HH size (people)",
      agea   = "Age (years)",
      gndr  = "Gender (0-female, 1-male)",
      year  = "Year"
    )
)
```

## Regression analysis

```{r}
fit1 <- 
  lm(
    stfeco ~ treat + y2010 + treat:y2010 + hhmmb + gndr + agea + 
      marr + agesq + urban + empl + sempl + other_work + cagriculture  + 
      cindustry + cfinance + factor(idregion),
    data = dta %>% zap_labels()
  )

modelsummary(
  list(fit1, fit1, fit1),
  statistic = NULL, estimate = "{estimate}{stars} ({std.error})",
  vcov = list(NULL,"robust", ~ idregion))

```


### Testing the assumptions

#### Visually

```{r}
check_model(fit1, check = c("linearity", "homogeneity"))
```

#### Varience homogeneity

```{r}
lmtest::bptest(fit1)
```

#### Collinearity

```{r}
check_collinearity(fit1)
```

