---
title: "Reproducing Card and Krueger 1994 paper"
author: "EB"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE,
  warning = FALSE
  )
```

# Loading and exploring data

```{r}
library(tidyverse)
library(GGally)
library(modelsummary)
library(performance)
library(lmtest)

dta <- read_rds("CardKrueger1994.rds") %>% 
  select(NJ, time, ncalls, inctime, firstinc, open, emptot, 
         pct_fte, wage_st, hrsopen, nregs, nregs11, bonus, chain, co_owned,
         observation, state, sheet)
glimpse(dta)
```

Present variables:

-   sheet - sheet number (unique store id) 
-   chain - chain 1=bk; 2=kfc; 3=roys; 4=wendys 
-   co_owned - 1 if company owned 
-   state - 1 if NJ; 0 if Pa 
-   ncalls - number of call-backs*
-   wage_st - starting wage ($/hr) 
-   inctime - months to usual first raise 
-   firstinc - usual amount of first raise ($/hr) 
-   bonus - 1 if cash bounty for new workers 
-   pctaff - % employees affected by new minimum 
-   meals - free/reduced price code 
-   open - hour of opening 
-   hrsopen - number hrs open per day 
-   psoda - price of medium soda, including tax 
-   pfry - price of small fries, including tax 
-   pentree - price of entree, including tax 
-   nregs - number of cash registers in store 
-   nregs11 - number of registers open at 11:00 am 
-   region - region of restaurant 
-   emptot - Total employment in a firm, FTE
-   pct_fte - Share of permanent employees, %
-   hrsopen - number hrs open per day


## Descriptive statistics

Summarize key variables 

```{r}
```

Perform t-test of means difference between groups by treatment status and the state.

```{r}
```

## DiD estimators

```{r}
```


