---
title: "Instrumental Variable"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

# Setup

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(haven)
library(modelsummary)
library(ivreg)
```

# Exercise 1. Reproduce Acemoglu et al. 2001

Acemoglu, Daron / Johnson, Simon / Robinson, James A. The Colonial Origins of Comparative Development: An Empirical Investigation. 2001. American Economic Review , Vol. 91, No. 5. p. 1369-1401

## 1 Data loading


In the data set, there are following variables:

-   `iso` - Country code;
-   `base_sample` - dummy variable indicating countries that constitute
    a base sample;
-   `africa`, `asia` and `other` - dummy variables indicating if a
    country is in Africa, Asia or on another continent. Please note that
    when all three dummy variables are equal to zero, that means that
    the country is in Latin America.
-   `pgp95` - GDP per capita (PPP) in 1995;
-   `hjypl` - Log output per worker in 1988;
-   `avexpr` - Average protection against expropriation risk,
    1985--1995;
-   `extmort4` - European settler mortality;
-   `lat_abst` - absolute value of the latitude, where the country is located;


```{r}
dta <-
  read_csv("Acemoglu2001.csv") %>% 
  filter(iso != "notIndonesia") %>% 
  mutate(l_pgp95 = log(pgp95), 
         l_extmort4 = log(extmort4))
```

-   filter out country that is called "notIndonesia";
-   compute natural logarithm of the variables, GDP per capita, output
    per worker and settlers mortality;

##  Descriptive statistics

```{r}
datasummary_skim(dta)
```

# OLS

```{r}
fit1 <- lm(l_pgp95 ~ avexpr, data = dta)
fit3 <- update(fit1, . ~ . + lat_abst)
fit5 <- update(fit3, . ~ . + asia + africa + other)

modelsummary::modelsummary(
  list(`(1) Whole world` = fit1, 
       `(3) Whole world` = fit3, 
       `(5) Whole world` = fit5
       ),
  fmt = "%.3f",
  estimate = "{estimate}{stars} ({std.error})",
  statistic = NULL,
  gof_map = c("adj.r.squared", "nobs"),
  stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  coef_rename = 
    c("avexpr" = "Av. prot. ag. expropriation risk 1985–1995",
      "lat_abst" = "Latitude",
      "asia" = "Asia dummy",
      "africa" = "Africa dummy",
      "other" = "Other continent dummy"), 
  coef_omit = "Intercep"
) 
```


# IV

```{r}
library(ivreg)
library(lmtest)

fit1iv <- ivreg(l_pgp95 ~ avexpr | l_extmort4,
                data = dta %>% filter(base_sample == 1),
                method = "M")

coef(fit1iv, component = "stage1")
vcov(fit1iv, component = "stage1") %>% diag %>% sqrt()

fit2iv <- update(fit1iv, . ~ . + lat_abst| . + lat_abst)

coef(fit2iv, component = "stage1")
vcov(fit2iv, component = "stage1") %>% diag %>% sqrt()

fit8iv <- update(fit1iv, . ~ . + lat_abst  + asia + africa + other| . + lat_abst  + asia + africa + other)

coef(fit2iv, component = "stage1")
vcov(fit2iv, component = "stage1") %>% diag %>% sqrt()

```

Second stage 

```{r}
modelsummary::modelsummary(
  list(`(1) Base sample` = fit1iv, 
       `(2) Base sample` = fit2iv,
       `(7) Base sample` = fit8iv
       ),
  fmt = "%.3f",
  estimate = "{estimate}{stars} ({std.error})",
  statistic = NULL,
  gof_map = c("adj.r.squared", "nobs", "statistic.Weak.instrument",
              "statistic.Wu.Hausman"),
  coef_rename = 
    c("avexpr" = "Av. prot. ag. expropriation risk 1985–1995",
      "lat_abst" = "Latitude",
      "asia" = "Asia dummy",
      "africa" = "Africa dummy",
      "other" = "Other continent dummy"), 
  stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  diagnostic = TRUE,
  title = "Panel A: Two-Stage Least Squares"
) 
```


First stage 

```{r}
list(
  `(1) Base sample` = fit1iv,
  `(2) Base sample` = fit2iv,
  `(7) Base sample` = fit8iv
) %>% 
  imap(~{
    modd <- .x
    moddname <- .y
    tibble(
      parameter = names(coef(modd, component = "stage1")),
      est = coef(modd, component = "stage1"),
      se = vcov(modd, component = "stage1") %>% diag() %>% sqrt()
    ) %>% 
      mutate(
        {{moddname}} := str_c(round(est, 3), " (", round(se, 3), ")")
      ) %>% 
      select(-se, -est)
  }) %>% 
  reduce(full_join) %>% 
  right_join(
    tibble(
      parameter = c("log_extmort4", "lat_abst", "asia", "africa", "other"),
      ` ` = c("Log European settler mortality", "Latitude", "Asia dummy",
              "Africa dummy", "Other continent dummy")
      )
  ) %>% 
  select(` `, everything()) %>% 
  select(-parameter ) %>% 
  knitr::kable()
```


