---
title: "Instrumental Variable"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

# Setup

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(haven)
library(modelsummary)

```

# Goals:

-   learn how to estimate an instrumental variable regression.

# Exercise 1. Reproduce Angrist & Krueger (1991)

Angrist, J. D., & Krueger, A. B. (1991). Does Compulsory School Attendance Affect Schooling and Earnings? The Quarterly Journal of Economics, 106, 979--1014. https://doi.org/10.2307/2937954

## 1 Data loading

Following variables are present:

-   `lnw` log if wage
-   `schooling`  years of schooling
-   `yob` year of birth
-   `qob` quarter of birth
-   `sob` week of birth

```{r}
dta <- 
  read_dta("ak91.dta") %>% 
  mutate(across(c(yob, qob, sob), ~ as.factor(.))) %>% 
  rename(schooling = s)
glimpse(dta)
datasummary_skim(dta)
datasummary_skim(dta, type = "categorical")
```

## 2 Data visualisation

```{r}
library(lubridate)
plot_dta <- 
  dta %>% 
  mutate(date = str_c("19", yob, "-", qob) %>% 
           lubridate::yq()) %>% 
  group_by(date) %>%
  summarise(across(c(lnw, schooling), ~mean(.))) %>% 
  mutate(yob = lubridate::year(date),
         qob = lubridate::quarter(date) %>% as_factor())
  
plot_dta %>% 
  ggplot() + 
  aes(y = lnw, x = date) + 
  geom_path() + 
  geom_label(aes(label = qob, fill = qob), colour = "white", 
             fontface = "bold", show.legend = FALSE) + 
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") + 
  xlab("Year of birth") + 
  ylab("Log of the weekly wage")

plot_dta %>% 
  ggplot() + 
  aes(y = schooling, x = date) + 
  geom_path() + 
  geom_label(aes(label = qob, fill = qob), colour = "white", fontface = "bold") + 
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") + 
  xlab("Year of birth") + 
  ylab("Years of schooling")
```


## 2 Produce IV using lm (only for the example)

Never use such manual 2sls estimations in real life as your SE are inefficient.

### 2.1 OLS

```{r}
fit1 <- lm(lnw ~ schooling + yob, dta)
summary(fit1)
```


### 2.2 first stage with qurters of birth

```{r}
fit_first_stage_qob1 <- lm(schooling ~ qob1 + yob, 
                      dta %>% mutate(qob1 = as.integer(qob == 1)))

fit_first_stage_qob1_noiv <- lm(schooling ~ yob, 
                                dta %>% mutate(qob1 = as.integer(qob == 1)))

fit_first_stage_allqob <-
  lm(schooling ~ -1 + qob1 + qob2 + qob3 + qob4 + yob, 
     dta %>% mutate(qob1 = as.integer(qob == 1),
                    qob2 = as.integer(qob == 2),
                    qob3 = as.integer(qob == 3),
                    qob4 = as.integer(qob == 4)))
```


```{r}
modelsummary(
  list(`First stage with IV` =  fit_first_stage_qob1,
       `First stage without the IV` =  fit_first_stage_qob1_noiv,
       `First stage with all birth quarters` =  fit_first_stage_allqob), 
  estimate = "{estimate} ({std.error}){stars}",
  statistic = NULL,
)
```


## 2.3 second stage with the qurter of birth

```{r}
fit_second_stage_qob1 <-
  lm(lnw ~ schooling_hat + yob, 
     dta %>% mutate(schooling_hat = fitted(fit_first_stage_qob1)))

fit_second_stage_allqob <-
  lm(lnw ~ schooling_hat + yob, 
     dta %>% mutate(schooling_hat = fitted(fit_first_stage_allqob)))

```

## 2.4 compare the models

```{r}
modelsummary(
  list(
    `OLS` = fit1,
    `Second stage QOB = 1` = fit_second_stage_qob1,
    `Second stage all QOB` = fit_second_stage_allqob
  ),
  fmt = "%.3f",
  estimate = "{estimate}{stars} ({std.error})",
  statistic = NULL,
  gof_map = c(
    "adj.r.squared",
    "nobs",
    "statistic.Weak.instrument",
    "statistic.Wu.Hausman",
    "F",
    "p.value"
  ),
  coef_rename =
    c(
      "schooling" = "Years of schooling",
      "schooling_hat" = "Years of schooling"
    ),
  coef_omit = "yob|Intercept",
  stars = c("*" = 0.05, "**" = 0.01, "***" = 0.001),
  diagnostic = TRUE,
  title = "Manual 2SLS"
)
```

## 3. Make an IV estimation

```{r}
library(ivreg)
iv_q1 <- ivreg(
  lnw  ~ schooling + yob | qob1 + yob,
  data = dta %>% mutate(qob1 = as.integer(qob == 1))
)

summary(iv_q1, diagnostics = T)

coef(iv_q1, component = c("stage1"), complete = FALSE)
coef(iv_q1, component = c("stage2"), complete = FALSE)


iv_qall <- ivreg(lnw  ~ schooling + yob | qob + yob,  data = dta )


coef(iv_q1, component = c("stage1"), complete = FALSE)
coef(iv_q1, component = c("stage2"), complete = FALSE)

summary(iv_qall, diagnostics = T)

```


```{r}

modelsummary(
  list(
    `OLS` = fit1,
    `Second stage QOB = 1` = fit_second_stage_qob1,
    `Second stage all QOB` = fit_second_stage_allqob,
    `Second stage QOB = 1` = iv_q1,
    `Second stage all QOB` = iv_qall
  ),
  fmt = "%.3f",
  estimate = "{estimate}{stars} ({std.error})",
  statistic = NULL,
  coef_omit = "yob|Intercept",
  diagnostic = TRUE,
  title = "Manual 2SLS"
)
```

